Semi-permeable vesicles: mathematical formulation, numerics and physics

### 2019-05-01 vesicle_code/docs/permeableVes/main.tex
fixed a few small typos and wordings

\section{Numerics}
To address stiffness associated with high-order derivatives,
equation~\eqref{eqn:vesVelocity} is discretized with a semi-implicit
time stepping method similar to our previous works.  First, we introduce
the notation
\begin{align}
  B[\mbx]\mbf &= -\kappa_b\frac{d^4}{ds^4} \mbf, \\
  T[\mbx]\sigma &= (\sigma \mbx_s)_s, \\
  P[\mbx]\mbf &= (\mbf \cdot \mbn) \mbn, \\
  D[\mbx]\mbf &= \mbx_s \cdot \mbf_s
\end{align}
where $\mbn$ is the is the outward unit normal of the vesicle
parameterized by $\mbx$, and $d/ds$ is its arclength derivative.
Dropping the dependence of the operators on $\mbx$, the dynamics of the
vesicle are governed by
\begin{align}
  \dot{\mbx} &= \mbu_{\infty}(\mbx) + 
  \beta P(B\mbx + T\sigma) + \sgl(B\mbx + T\sigma), \\
  D \dot{\mbx} &= 0.
\end{align}
Using the notation $B^N$ to denote the bending operator due to the
vesicle configuration at time $t_N$, and similar notation for the other
operators, we apply the semi-implicit time stepping method
\begin{align}  
  \frac{\mbx^{N+1} - \mbx^N}{\Delta t} = \mbu_\infty(\mbx^N) 
  + \beta P^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}) 
  + \sgl^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}).
\end{align}
Applying $D$ to the first-order time discretization of
equation~\eqref{eqn:vesVelocity}, the inextensibility condition is
\begin{align*}
  &D^N(\mbx^{N+1} - \mbx^N) = 0 \\ 
  \Rightarrow &D^N\mbx^{N+1} = D^N \mbx^N \\ 
  \Rightarrow &D^N\mbx^{N+1} = 1.
\end{align*}
Finally, we write the time stepping method in matrix form as
\begin{align}
  A \left(
    \begin{array}{c}
      \mbx^{N+1} \\ \sigma^{N+1}
    \end{array}
  \right) = 
  \left(
    \begin{array}{c}
      \mbx^{N} + \Delta t \mbu_\infty(\mbx^N) \\ \mathbf{1}
    \end{array}
  \right),
\end{align}
where
\begin{align}
 A = \left(
  \begin{array}{cc}
    I + \Delta t \beta P^N B^N + \Delta t \sgl^N B^N & 
    -\Delta t P^N T^N - \Delta t \sgl^N T^N \\
    D^N & 0
  \end{array}
  \right).
\end{align}

\section{Including a concentration field}
Based on the work of Yao and Mori~\cite{yao-mor2017}, we introduce a
chemical concentration field $c$.  The concentration field is governed
by the advection-diffusion equation
\begin{align}
  \pderiv{c}{t} + \nabla \cdot (\mbu c) = D\Delta c.
\end{align}
If the concentration is diffusion dominated (zero Peclet number), then
the concentration satisfies
\begin{align}
  \Delta c = 0, \quad \mbx \in \RR^2.
\end{align}
The boundary conditions of the concentration field on the vesicle
$\gamma$ are
\begin{align}
  (\mbu c - D\nabla c) \cdot \mbn = c \pderiv{\mbx}{t} \cdot \mbn + 
    j_c + j_p,
\end{align}
where $j_c$ and $j_p$ are the transmembrane chemical flux due to passive
transport and active transport (pumps), respectively.  In the zero
Peclet number limit, the absence of an active transport mechanism, and
the choice for the passive flux, the boundary condition of the
concentration field is
\begin{align}
  -D\pderiv{c}{\mbn} = c \pderiv{\mbx}{t} + 
    k_c \jump{c}, \quad \mbx \in \gamma.
\end{align}
To couple the vesicle dynamics to the concentration, the vesicle
velocity satisfies the slip condition
\begin{align}
  \mbu - \pderiv{\mbx}{t} = j_w \mbn,
\end{align}
where $j_w$ is the water flux through the vesicle.
In~\cite{yao-mor2017}, this flux is
\begin{align}
  j_w = -k_w(RT\jump{c} + \mbf_\text{mem} \cdot \mbn).
\end{align}
In our current implementation, we drop the $RT\jump{c}$ term, and
let $\beta = k_w$.

@@

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/1.jpg}
	\includegraphics[width=.9\textwidth]{figures/2.jpg}
	\includegraphics[width=.9\textwidth]{figures/3.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
	\label{Shear0}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/4.jpg}
	\includegraphics[width=.9\textwidth]{figures/5.jpg}
	\includegraphics[width=.9\textwidth]{figures/6.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
	\label{Shear0}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/7.jpg}
	\includegraphics[width=.9\textwidth]{figures/8.jpg}
	\includegraphics[width=.9\textwidth]{figures/9.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
	\label{Shear0}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/10.jpg}
	\includegraphics[width=.9\textwidth]{figures/11.jpg}
	\includegraphics[width=.9\textwidth]{figures/12.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 2$.}
	\label{Shear0}
\end{figure}

 \begin{figure}
 	\centering
 	\includegraphics[width=.9\textwidth]{figures/BE1.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE2.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE3.jpg}
 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
 	\label{Shear0}
 \end{figure}

 \begin{figure}
 	\centering
 	\includegraphics[width=.9\textwidth]{figures/BE4.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE5.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE6.jpg}
 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
 	\label{Shear0}
 \end{figure}

 \begin{figure}
 	\centering
 	\includegraphics[width=.9\textwidth]{figures/BE7.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE8.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE9.jpg}
 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
 	\label{Shear0}
 \end{figure}

>>>>

\section{Numerics}
To address stiffness associated with high-order derivatives,
equation~\eqref{eqn:vesVelocity} is discretized with a semi-implicit
time stepping method similar to our previous works.  First, we introduce
the notation
\begin{align}
  B[\mbx]\mbf &= -\kappa_b\frac{d^4}{ds^4} \mbf, \\
  T[\mbx]\sigma &= (\sigma \mbx_s)_s, \\
  P[\mbx]\mbf &= (\mbf \cdot \mbn) \mbn, \\
  D[\mbx]\mbf &= \mbx_s \cdot \mbf_s,
\end{align}
where $\mbn$ is the outward unit normal of the vesicle parameterized by
$\mbx$, and $d/ds$ is its arclength derivative.  Dropping the dependence
of the operators on $\mbx$, the dynamics of the vesicle are governed by
\begin{align}
  \dot{\mbx} &= \mbu_{\infty}(\mbx) + 
  \beta P(B\mbx + T\sigma) + \sgl(B\mbx + T\sigma), \\
  D \dot{\mbx} &= 0.
\end{align}
Using the notation $B^N$ to denote the bending operator due to the
vesicle configuration at time $t_N$, and similar notation for the other
operators, we apply the semi-implicit time stepping method
\begin{align}  
  \frac{\mbx^{N+1} - \mbx^N}{\Delta t} = \mbu_\infty(\mbx^N) 
  + \beta P^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}) 
  + \sgl^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}).
\end{align}
Applying $D$ to the first-order time discretization of
equation~\eqref{eqn:vesVelocity}, the inextensibility condition is
\begin{align*}
  &D^N(\mbx^{N+1} - \mbx^N) = 0 \\ 
  \Rightarrow &D^N\mbx^{N+1} = D^N \mbx^N \\ 
  \Rightarrow &D^N\mbx^{N+1} = 1.
\end{align*}
Finally, we write the time stepping method in matrix form as
\begin{align}
  A \left(
    \begin{array}{c}
      \mbx^{N+1} \\ \sigma^{N+1}
    \end{array}
  \right) = 
  \left(
    \begin{array}{c}
      \mbx^{N} + \Delta t \mbu_\infty(\mbx^N) \\ \mathbf{1}
    \end{array}
  \right),
\end{align}
where
\begin{align}
 A = \left(
  \begin{array}{cc}
    I + \Delta t \beta P^N B^N + \Delta t \sgl^N B^N & 
    -\Delta t P^N T^N - \Delta t \sgl^N T^N \\
    D^N & 0
  \end{array}
  \right).
\end{align}

\section{Including a concentration field}
Based on the work of Yao and Mori~\cite{yao-mor2017}, we introduce a
chemical concentration field $c$.  The concentration field is governed
by the advection-diffusion equation
\begin{align}
  \pderiv{c}{t} + \nabla \cdot (\mbu c) = D\Delta c.
\end{align}
If the concentration is diffusion dominated (zero Peclet number), then
the concentration satisfies
\begin{align}
  \Delta c = 0, \quad \mbx \in \RR^2.
\end{align}
The boundary conditions of the concentration field on the vesicle
$\gamma$ are
\begin{align}
  (\mbu c - D\nabla c) \cdot \mbn = c \pderiv{\mbx}{t} \cdot \mbn + 
    j_c + j_p,
\end{align}
where $j_c$ and $j_p$ are the transmembrane chemical flux due to passive
transport and active transport (pumps), respectively.  In the zero
Peclet number limit, the absence of an active transport mechanism, and
the choice for the passive flux, the boundary condition of the
concentration field is
\begin{align}
  -D\pderiv{c}{\mbn} = c \pderiv{\mbx}{t} + 
    k_c \jump{c}, \quad \mbx \in \gamma.
\end{align}
To couple the vesicle dynamics to the concentration, there is a
difference between the fluid velocity and vesicle velocity
\begin{align}
  \mbu - \pderiv{\mbx}{t} = j_w \mbn,
\end{align}
where $j_w$ is the water flux through the vesicle.  This is the
osmophoretic flow.  In~\cite{yao-mor2017}, this flux is
\begin{align}
  j_w = -k_w(RT\jump{c} + \mbf_\text{mem} \cdot \mbn).
\end{align}
In our current implementation, we drop the $RT\jump{c}$ term, and
let $\beta = k_w$.

@@

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/1.jpg}
	\includegraphics[width=.9\textwidth]{figures/2.jpg}
	\includegraphics[width=.9\textwidth]{figures/3.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
%	\label{Shear0}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/4.jpg}
	\includegraphics[width=.9\textwidth]{figures/5.jpg}
	\includegraphics[width=.9\textwidth]{figures/6.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
%	\label{Shear0}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/7.jpg}
	\includegraphics[width=.9\textwidth]{figures/8.jpg}
	\includegraphics[width=.9\textwidth]{figures/9.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
%	\label{Shear0}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.9\textwidth]{figures/10.jpg}
	\includegraphics[width=.9\textwidth]{figures/11.jpg}
	\includegraphics[width=.9\textwidth]{figures/12.jpg}
	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 2$.}
%	\label{Shear0}
\end{figure}

 \begin{figure}
 	\centering
 	\includegraphics[width=.9\textwidth]{figures/BE1.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE2.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE3.jpg}
 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
% 	\label{Shear0}
 \end{figure}

 \begin{figure}
 	\centering
 	\includegraphics[width=.9\textwidth]{figures/BE4.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE5.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE6.jpg}
 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
% 	\label{Shear0}
 \end{figure}

 \begin{figure}
 	\centering
 	\includegraphics[width=.9\textwidth]{figures/BE7.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE8.jpg}
 	\includegraphics[width=.9\textwidth]{figures/BE9.jpg}
 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
% 	\label{Shear0}
 \end{figure}
